Value Ordering, Discrepancies, and Restarts for Subgraph Algorithms\thanks{This work was supported by the Engineering and Physical Sciences Research Council [grant numbers EP/026842/1 and EP/M508056/1]}

### 2018-01-18 paper/paper.tex
Things
Patrick's changes

\begin{abstract}
    Modern subgraph isomorphism solvers use degree-based heuristics to direct a constraint
    programming style search. We show that discrepancy-based searches are too expensive to be
    effective in offsetting mistakes made by value-ordering heuristic in this setting.  We then
    demonstrate that a random heuristic with restarts beats conventional search with the best-known
    tailored heuristic. This motivates the introduction of a slightly-random heuristic with similar
    behaviour to the tailored heuristic: when combined with restarts, this new search strategy is
    over a hundred times more effective on satisfiable instances, and does not affect performance on
    unsatisfiable instances. Finally, we show that the same strategy also improves two maximum
    common induced subgraph algorithms.
\end{abstract}

@@

Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself. The
\emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood; when the graph is
clear from the context, we omit the $G$ subscript. The \emph{neighbourhood degree sequence} of $v$,
$\nds_G(v)$, is the sequence of degrees of the neighbours of $v$ in descending order, and we denote
by $\preceq$ the inclusion relation of \citet{DBLP:journals/constraints/ZampelliDS10}. As per
\citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$,
and with edges between vertices $v$ and $v$ if there are at least $n$ simple paths of length exactly
$\ell$ between vertices $v$ and $w$ in $G$.

@@

The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
(\siplineref{variableordering}), and tries assigning
it each of its remaining value in turn (\siplineref{valueordering}). The effects of this assignment
are propagated (\siplineref{dopropagate}). If an
inconsistency is detected, we continue and try another value; otherwise, we recurse
(\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
(\siplineref{backtrack}).

@@

\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
current state of the art.

@@

The performance of the degree and biased heuristics are very similar: the
biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
performance above it. This shows that we can introduce an element of randomness into the degree
value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
shows that despite the aggregate performance being very similar, on an case by case basis, the two
heuristics can make a large difference to the performance of satisfiable instances.

@@

>>>>

\begin{abstract}
    Modern subgraph isomorphism solvers use degree-based heuristics to direct a constraint
    programming style search. We show that discrepancy-based searches are too expensive to be
    effective in offsetting mistakes made by value-ordering heuristics in this situation.  We then
    demonstrate that a random heuristic with restarts beats conventional search with the best-known
    tailored heuristic. This motivates the introduction of a slightly-random heuristic with similar
    behaviour to the tailored heuristic: when combined with restarts, this new search strategy is
    over a hundred times more effective on satisfiable instances, and does not affect performance on
    unsatisfiable instances. Finally, we show that the same strategy also improves two maximum
    common induced subgraph algorithms.
\end{abstract}

@@

Let $G$ be a graph. We denote its set of vertices by $\vertexset(G)$, and write $v \sim_G w$ to mean
that vertices $v$ and $w$ are adjacent; we allow a vertex to be adjacent to itself. The
\emph{neighbourhood} of a vertex, $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$, and
the \emph{degree} of $v$, $\deg_G(v)$, is the cardinality of its neighbourhood; when the graph is
clear from the context, we omit the $G$ subscript. The \emph{neighbourhood degree sequence} of $v$,
$\nds_G(v)$, is the sequence of degrees of the neighbours of $v$ in descending order, and we denote
by $\preceq$ the inclusion relation of \citet{DBLP:journals/constraints/ZampelliDS10}. As per
\citet{DBLP:conf/aaai/HoffmannMR17}, $G^{n,\ell}$ denotes the graph with vertex set $\vertexset(G)$,
and with edges between vertices $v$ and $w$ if there are at least $n$ simple paths of length exactly
$\ell$ between vertices $v$ and $w$ in $G$.

@@

The $\FuncSty{search}$ procedure recursively selects an unassigned domain $D$
(\siplineref{variableordering}), and tries assigning
it each of its remaining values in turn (\siplineref{valueordering}). The effects of this assignment
are propagated (\siplineref{dopropagate}). If an
inconsistency is detected, we continue and try another value; otherwise, we recurse
(\siplineref{recurse}) until a solution is found. Finally, if we run out of values, we backtrack
(\siplineref{backtrack}).

@@

\paragraph{Degree} We select vertices from highest degree to lowest degree. This heuristic is the
current state of the art \citep{DBLP:conf/ijcai/McCreeshPT16}.

@@

The performance of the degree and biased heuristics are very similar: the
biased heuristic has slightly worse performance below a 400s timeout, and effectively equal
performance above it. This shows that we can introduce an element of randomness into the degree
value-ordering heuristic without adversely affecting its performance \emph{in aggregate}.
\Cref{figure:scatter-heuristics} gives a more detailed look at these results. This scatter plot
shows that despite the aggregate performance being very similar, on a case by case basis, the two
heuristics can make a large difference to the performance of satisfiable instances.

@@

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-constant.pdf}

    \caption{Constant restarts?}
    \label{figure:scatter-constant}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-parallel.pdf}

    \caption{Parallel}
    \label{figure:parallel}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-parallel.pdf}

    \caption{Parallel Luby}
    \label{figure:parallel-luby}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics*{gen-graph-scatter-parallel-constant.pdf}

    \caption{Parallel Constant}
    \label{figure:parallel-constant}
\end{figure}
